### 2016-03-09 Projects/DelCo_2014_2015/Reports/Hose_Stream_Report/TeX_Files/DelCo_Hose_Stream_Tests.tex
Updated hose stream report, deleted unnecessary figs

Fire suppression tactics using hose streams also affect the ventilation in a structure and can impact the movement of smoke and heat through a structure as vents are made to advance the hose line or if ventilation inducing handline tactics are in practice~\cite{?}. Additional research addressing the coordination of suppression tactics and the impact on ventilation is needed to complete recommendations on fire control tactics to appropriate standards, education, and training documents.

There exists a multitude of different tools, methods, and techniques used to apply water for fire suppression using hose streams. A common type of nozzle that is used to apply water during fire suppression in residential structures is a fog nozzle. Most fog nozzles contain an adjustable tip that can be rotated to produce different stream patterns and are therefore often referred to as combination nozzles. The types of streams produced by a fog nozzle are often divided into three categories based on the angle of the stream with respect to the nozzle: a straight stream typically produces an angle less than 15\SI{}{\degree}, a narrow fog stream produces an angle between 15\SI{}{\degree} and 45\SI{}{\degree}, and a wide fog stream produces an angle between 45\SI{}{\degree} to 80\SI{}{\degree}~\cite{IFSTA:Essentials_of_FF} \textit{(cited 5th edition because that's where I got this info from, I'm assuming I can change it to 6th?)}. Increasing the angle of the stream allows for a larger area to be covered by water, thus exposing a larger water droplet surface area for heat absorption. However, increasing the angle decreases the distance reached by the stream and its penetration power and increases the potential to disrupt thermal layering increases. 
%\textit{Essentials of Fire Fighting} defines 
%https://docs.google.com/presentation/d/1hZ5DZhYXxCkcIEHNZKumH48b_Q53iM2D7asXkMYlVMM/embed?hl=en&size=l&slide=id.p39

%http://www.tft.com/literature/library/files/bfa.pdf
%Keith Royer, "Water for Fire Fighting", (Engineering Extension Service, Iowa State University, Ames, Iowa) p. 6 
In addition to the type of stream pattern used, fire suppression tactics using hose streams can differ in the method used to apply the hose stream. For example, Royer suggests using a clockwise rotation instead of a counterclockwise application pattern for the following reasons~\cite{Royer:ISU}:
\begin{enumerate} 
	\item Clockwise rotation drives most of the heated gases, smoke, and flames away from the nozzle. Counterclockwise rotation does just the opposite.
	\item Steam formed by the clockwise rotation has a violent rolling action. Counterclockwise rotation produces steam with an inactive and lazy action.
	\item A clockweise roation increases the efficiency of water and produces a faster knockdown time than counterclockwise rotation.
\end{enumerate}

\section{Objectives}
\label{sec:objectives}
This report is concerned with examining the impact of hose stream and application pattern selection on air movement inside a residential structure. Numerous experiments containing different flow path configurations were conducted to study the impact of hose stream and water application selection on ventilation in a structure. More specifically, the experiments were conducted to fulfill the following objectives:
\begin{enumerate}
	\item Determine which hose stream patterns cause significant air movement in a structure 
	\item Determine the amount of air movement induced by different stream patterns
	\item Determine the impact of different stream application patterns on the ventilation in a structure
\end{enumerate}

Three types of hose streams and four different water application patterns were studied throughout the experiments described in this report. The three types of hose stream patterns studied, pictured in Figure~\ref{fig:hose_streams}, were straight stream, narrow fog stream, and wide fog stream. Water was flowed into the structure by having the hose in a static, or fixed, position, by moving the hose line left-to-right across the room in a sweeping motion, and by rotating the hose line in both the clockwise and counterclockwise directions.

% \begin{figure}[!ht]
% 	\includegraphics[width=\columnwidth]{../Figures/Pictures/hose_streams.pdf}
% 	\caption[Hose stream patterns used in experiments.]{The three types of stream patterns used during water flow experiments. From left to right: straight stream, narrow fog stream, and wide fog stream.}
% 	\label{fig:hose_streams}
% \end{figure}

\begin{figure}[!ht]
	\includegraphics[width=0.55\columnwidth]{../Figures/Pictures/hose_streams_1}
	\\~\\
	\includegraphics[width=0.55\columnwidth]{../Figures/Pictures/hose_streams_2}
	\caption[Hose stream patterns used in experiments.]{The four types of stream patterns used during water flow experiments. Top row: narrow fog stream (left) and wide fog stream (right). Bottom row: straight stream from combination nozzle (left) and solid stream from smooth bore nozzle (right).}
	\label{fig:hose_streams}
\end{figure}

Additionally, a set of experiments comparing the impact of a straight stream from a combination nozzle and a solid stream from a smooth bore nozzle on air flows in a structure was conducted.

@@

The interior walls of the first floor of each structure were framed with steel studs set to 0.40~m (16~in) centers and track and were lined with 13~mm (0.5~in) thick cement board. The walls were composed of 16~mm (5/8~in) Type X gypsum board. Additionally, the ceiling was composed of two layers of 13~mm (0.5~in) thick cement board.
\FloatBarrier

The first floor ceiling support of each structure was composed of wood truss joist I-beams (TJIs) with a 299~mm (11.75~in) depth. Each TJI was composed of laminated veneer lumber flanges with a cross section of 29~mm (1.13~in) x 44~mm (1.75~in) and an 11~mm (0.43~in) thick oriented strand board web as shown in Fig.~\ref{fig:TJI}. Tongue and grove, 18.3~mm (0.72~in) thick, oriented strand board was screwed to the top of the TJIs.

@@

The second floor of the West Structure was built on the wood ceiling support described above and was connected to the first floor by a stairwell. The second story walls were wood framed with 51~mm (2~in) by 102~mm (4~in) studs set to 0.40~m (16~in) centers. The interior walls were protected by 16~mm (5/8~in) fire rated gypsum board, 16~mm (5/8~in) durarock board, and a second layer of 16~mm (5/8~in) fire rated gypsum board. The exterior walls were protected with 11~mm (7/16~in) oriented strand board and 8~mm (5/16~in) fiber cement lap siding.

\subsubsection{Layout}
\label{sec:layout}
Dimensioned floor plans of the East and West Structures are presented in Figures~\ref{fig:east_dimensioned_plan}~and~\ref{fig:west_dimensioned_plan}, respectively. \textbf{Add N arrow to floor plans}

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/East_Test_Structure_Dimensioned_Full}
	\caption[Dimensioned floor plan of the East Structure.]{East Structure floor layout. Interior dimensions are symmetric across horizontal and vertical centerlines.}
	\label{fig:east_dimensioned_plan}
\end{figure}

@@

\subsection{Instrumentation}
\label{sec:instrumentation}
A discussion of uncertainties for each type of measurement in this report can be found in Section~\ref{sec:uncertainty}. For every experiment, gas velocity at an interior doorway was measured using differential pressure transducers connected to bi-directional velocity probes~\cite{McCaffrey:Combustion_and_Flame}. An array of 6 bi-directional probes located at the northeast interior doorway of the East Structure is pictured below in Fig.~\ref{fig:BDPs}. A single thermocouple was attached to each bi-directional probes to measure the temperature to be used in conjunction with the pressure measurement to derive velocity. Every thermocouple attached to the bi-directional probes was exposed-bead, Chromel-Alumel (type K) with a 1.0~mm (0.04 in) diameter. Starting with the exposed bead, the thermocouple wire was sheathed in a 3.2~mm (0.13 in) diameter Inconel shield that was 0.76~m (2.5~ft) in length.

The approximate locations of the bi-directional probes used to measure air velocity are annotated in the plan views of the East Structure and West Structure experimental setups (Fig.~\ref{fig:east_setup} and Figs.~\ref{fig:flow_path_1}-\ref{fig:flow_path_2}, respectively) found in Section~\ref{sec:exp_procedure}. To minimize potential measurement errors due to external environmental conditions (i.e., wind), only velocity data measured by bi-directional probes located at interior doorways are considered in this report. The set of bi-directional probes at the northeast interior doorway of the East Structure, denoted as ``A6", contained 6 probes located at distances of 0.30 m, 0.60 m, 0.90 m, 1.20 m, 1.50 m, and 1.80~m below the soffit of the doorway. The set of bi-directional probes at the interior doorway in the West Structure, denoted as ``A10", contained 8 probes located at distances of 0.08 m, 0.34 m, 0.61 m, 0.88 m, 1.15 m, 1.42 m, 1.68 m, and 1.95~m below the soffit of the corresponding doorway.

\begin{figure}[!ht]
	\includegraphics[width=3.5in]{../Figures/Pictures/BDPs_east}
	\caption[Array of bi-directional probes at interior doorway in East Structure.]{Array of bi-directional probes located at the northeast interior doorway in the East Structure. View is looking at the array from Room C.}
	\label{fig:BDPs}
\end{figure}
\FloatBarrier

@@

Water flow rate was measured with a pressure and flow meter combination shown in Fig.~\ref{fig:flow_meter}. The meter consists of a section of 6.35 cm (2.5 in) cast aluminum pipe with a 0-4.1 MPa (0-600 psi) pressure transducer and a paddlewheel type flow sensor with a range of 0 to 4800 lpm (1250 gpm). The pressure transducer and paddlewheel both connect to the battery operated control box where the pressure transducer voltage is converted to a pressure and the paddlewheel pulse count is converted to a volumetric flow rate. The manufacturer reports a $\pm5$\% calibration expanded uncertainty for the flow sensor and $\pm3$\%  for the pressure sensor~\cite{Akron}. The pressure transducer was calibrated with a known analog pressure gauge. The flow meter was calibrated by capturing water over time and measuring that mass of water to determine the flow rate. The total expanded uncertainty was estimated at $\pm10$\%.

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Pictures/flow_meter}
	\caption[Flow meter used to measure flow rate during experiments.]{Pressure and flow meter combination used to measure water flow rate during the experiments.}
	\label{fig:flow_meter}
\end{figure}
\FloatBarrier

\section{Experimental Procedure}
\label{sec:exp_procedure}
Every experiment discussed in this report involved discharging water at a known flow rate from a combination nozzle set to flow water in a straight, narrow fog, or wide fog stream pattern or from a smooth bore nozzle in a solid stream pattern. Specific doors were opened or closed before and/or during the experiments to change the ventilation in the structure.

A total of seven test series were conducted to study the impact of different hose stream and application patterns on gas movement in a structure; three used a combination nozzle attached to a monitor to flow water at 120~GPM, three used a combination nozzle attached directly to a hoseline to flow water at 120~GPM (Fig.~\ref{fig:monitor+handline}), and one used combination and smooth bore nozzles attached to a monitor to flow water at 180~GPM. Experiments that used a monitor to flow water always applied the stream in a fixed position and thus were focused on changing the location of water application in addition to the stream pattern, while the experiments that used a combination nozzle attached directly to a hoseline to flow water focused on varying the pattern in which water was applied in addition to the stream pattern.

\begin{figure}[!ht]
	\minipage{3in}
	\begin{center}
		\includegraphics[width=2.9in]{../Figures/Pictures/monitor}
	\end{center} 
	\endminipage \hfill
	\minipage{3in}
	\begin{center}
		\includegraphics[width=2.9in]{../Figures/Pictures/handline}
	\end{center}
	\endminipage \hfill
	\caption[Monitor and handline equipped with a combination nozzle.]{Monitor (left) and 1.75~in handline (right) equipped with a combination nozzle and used to flow water during experiments.}
	\label{fig:monitor+handline}
\end{figure}
\FloatBarrier

Three different layouts were used for the seven test series. The test series conducted in the East Structure, Tests 33 and 34, were performed using an identical structure layout, presented in Fig.~\ref{fig:east_setup}. Test 33 used a monitor and combination nozzle to flow water and Test 34 used a combination nozzle attached directly to a handline to flow water.

The main difference between the two layouts used for the West Structure experiments is the position of the south side door on the first level. Tests 16 and 18 used the configuration with the door opened (Fig.~\ref{fig:flow_path_1}) and Tests 17, 19, and 70 used the configuration with the door closed (Fig.~\ref{fig:flow_path_2}). Tests 16 and 17 used a combination nozzle attached to a monitor to flow water; Tests 18 and 19 used a combination nozzle attached directly to a handline to flow water; and Test 70 used a combination nozzle and a smooth bore nozzle attached to a monitor to flow water in a straight and solid stream, respectively.

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/Specific_Tests/East_Hose_Test_Annotated}
	\caption[Plan view of the East Structure setup for Tests 33 and 34.]{Plan view of the East Structure setup for Tests 33 and 34. The view is annotated with the approximate location of water flow (nozzle graphic), the direction of the established flow path (red line), and the approximate location of A6 (blue square).}
	\label{fig:east_setup}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/Specific_Tests/West_Hose_Test_2nd_Floor_Annotated}
	\\~\\
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/Specific_Tests/West_Hose_Test_18_1st_Floor_Annotated}
	\caption[Plan view of the West Structure setup for Tests 16 and 18.]{Plan view of the second floor (top) and first floor (bottom) setup in the West Structure for Tests 16 and 18. The approximate location of water flow (nozzle graphic), the direction of the established flow path (red line), and the approximate location of A10 (blue square). On the second floor, the stairwell door and north side, west double door were opened and closed at certain instances during the tests; the flow path was considered fully established when both doors were in the opened position.}
	\label{fig:flow_path_1}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/Specific_Tests/West_Hose_Test_2nd_Floor_Annotated}
	\\~\\
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/Specific_Tests/West_Hose_Test_19_1st_Floor_Annotated}
	\caption[Plan view of the West Structure setup for Tests 17 and 19.]{Plan view of the second floor (top) and first floor (bottom) setup in the West Structure for Tests 17 and 19. The approximate location of water flow (nozzle graphic), the direction of the established flow path (red line), and the approximate location of A10 (blue square). On the second floor, the stairwell door and north side, west double door were opened and closed at certain instances during the tests; the flow path was considered fully established when both doors were in the opened position.}
	\label{fig:flow_path_2}
\end{figure}
\FloatBarrier

\subsection{Monitor Experiments}
\label{sec:monitor_procedure}
Numerous experiments were conducted during three different test series using a monitor equipped with a combination nozzle to flow water at 120~GPM in a straight stream, narrow fog stream, and wide fog stream pattern. One test series, Test 33, occurred in the East Structure, and the two remaining series, Tests 16 and 17, occurred in the West Structure. An additional test series, Test 70, used a combination nozzle and a smooth bore nozzle attached to a monitor to flow water at 180~GPM in a straight stream and solid stream, respectively. Because monitor nozzles are designed to flow water from a set position, water was always applied in a fixed position during the monitor experiments. Thus, the monitor experiments primarily focused on the impact of different hose stream patterns on air movement in a structure.

Test 33 utilized a monitor equipped with a combination nozzle to flow water from Room C of the East Structure to the area above the doorway on the south wall of Room B (Fig.~\ref{fig:test_33_pic}). A plan view of the structure's layout annotated with the approximate location of the bi-directional probes array used to measure gas velocity through an interior doorway (A6) and the general direction of the fully established flow path for Test 33 can be found in Fig.~\ref{fig:east_setup}. The procedure for Test 33 was the simplest of all the test procedures. First, water flowed in a straight stream pattern for 30 seconds, then the stream was changed to a narrow fog for 30 seconds, and finally to a wide fog stream for 30 seconds. An identical process was repeated an additional time.

>>>>

Fire suppression tactics using hose streams also affect the ventilation in a structure and can impact the movement of smoke and heat through a structure as vents are made to advance the hose line or if ventilation inducing handline tactics are in practice. Additional research addressing the coordination of suppression tactics and the impact on ventilation is needed to complete recommendations on fire control tactics to appropriate standards, education, and training documents.

There exists a multitude of different tools, methods, and techniques used to apply water for fire suppression using hose streams. A common type of nozzle that is used to apply water during fire suppression in residential structures is a fog nozzle. Most fog nozzles contain an adjustable tip that can be rotated to produce different stream patterns and are therefore often referred to as combination nozzles. The types of streams produced by a fog nozzle are often divided into three categories based on the angle of the stream with respect to the nozzle: a straight stream typically produces an angle less than 15\SI{}{\degree}, a narrow fog stream produces an angle between 15\SI{}{\degree} and 45\SI{}{\degree}, and a wide fog stream produces an angle between 45\SI{}{\degree} to 80\SI{}{\degree}~\cite{IFSTA:Essentials_of_FF} \textit{(cited 5th edition because that's where I got this info from, I'm assuming I can change it to 6th?)}. Increasing the angle of the stream allows for a larger area to be covered by water, thus exposing a larger water droplet surface area for heat absorption. However, increasing the angle decreases the distance reached by the stream and its penetration power and increases the potential to disrupt thermal layering. 
%\textit{Essentials of Fire Fighting} defines 
%https://docs.google.com/presentation/d/1hZ5DZhYXxCkcIEHNZKumH48b_Q53iM2D7asXkMYlVMM/embed?hl=en&size=l&slide=id.p39

%http://www.tft.com/literature/library/files/bfa.pdf
%Keith Royer, "Water for Fire Fighting", (Engineering Extension Service, Iowa State University, Ames, Iowa) p. 6 
In addition to the type of stream pattern used, the method used to apply the hose stream can also vary. For example, Royer suggests using a clockwise rotation instead of a counterclockwise application pattern for the following reasons~\cite{Royer:ISU}:
\begin{enumerate} 
	\item Clockwise rotation drives most of the heated gases, smoke, and flames away from the nozzle. Counterclockwise rotation does just the opposite.
	\item Steam formed by the clockwise rotation has a violent rolling action. Counterclockwise rotation produces steam with an inactive and lazy action.
	\item A clockwise roation increases the efficiency of water and produces a faster knockdown time than counterclockwise rotation.
\end{enumerate}

\section{Objectives}
\label{sec:objectives}
This report is concerned with examining the impact of hose stream and application pattern selection on air movement inside a residential structure. Numerous experiments containing different flow path configurations were conducted to study the impact of hose stream and water application selection on ventilation in a structure. More specifically, the experiments were conducted to fulfill the following objectives:
\begin{enumerate}
	\item Determine which hose stream patterns cause significant air movement in a structure 
	\item Determine the amount of air movement induced by different stream patterns
	\item Determine the impact of different stream application patterns on the ventilation in a structure
\end{enumerate}
Four types of hose streams and four different water application patterns were studied throughout the experiments described in this report. Three types of hose stream patterns using a combination nozzle were studied: straight stream, narrow fog stream, and wide fog stream patterns. Water was flowed into the structure by having the hose in a static, or fixed, position, by moving the hose line left-to-right across the room in a sweeping motion, and by rotating the hose line in both the clockwise and counterclockwise directions. Additionally, a set of experiments comparing the impact of a straight stream from a combination nozzle and a solid stream from a smooth bore nozzle on air flows in a structure was conducted. Images of the four types of hose streams are presented in Fig.~\ref{fig:hose_streams}.

\begin{figure}[!ht]
	\begin{subfigure}[b]{0.45\columnwidth}
		\centering
		\includegraphics[width=0.75\columnwidth]{../Figures/Pictures/NF_example}
		\caption{Narrow fog stream}
	\end{subfigure}
	\begin{subfigure}[b]{0.45\columnwidth}
		\centering
		\includegraphics[width=0.75\columnwidth]{../Figures/Pictures/WF_example}
		\caption{Wide fog stream}
	\end{subfigure}
	\\~\\
	\begin{subfigure}[b]{0.45\columnwidth}
		\centering
		\includegraphics[width=0.75\columnwidth]{../Figures/Pictures/SS_70}
		\caption{Straight stream from combination nozzle}
	\end{subfigure}
	\begin{subfigure}[b]{0.45\columnwidth}
		\centering
		\includegraphics[width=0.75\columnwidth]{../Figures/Pictures/SB_70}
		\caption{Solid stream from smooth bore nozzle}
	\end{subfigure}
	\caption[Hose stream patterns used during experiments.]{The four types of hose stream patterns used during water flow experiments.}
	\label{fig:hose_streams}
\end{figure}

@@

The interior walls of the first floor of each structure were framed with steel studs set to 0.40~m (16~in) centers and track and were lined with 13~mm (0.5~in) thick cement board. The walls were composed of 16~mm (0.625~in) Type X gypsum board. Additionally, the ceiling was composed of two layers of 13~mm (0.5~in) thick cement board.
\FloatBarrier

The first floor ceiling support of each structure was composed of wood truss joist I-beams (TJIs) with a 299~mm (11.75~in) depth. Each TJI was composed of laminated veneer lumber flanges with a cross section of 29~mm (1.125~in) x 44~mm (1.75~in) and an 11~mm (0.43~in) thick oriented strand board web as shown in Fig.~\ref{fig:TJI}. Tongue and grove, 18.3~mm (0.72~in) thick, oriented strand board was screwed to the top of the TJIs.

@@

The second floor of the West Structure was built on the wood ceiling support described above and was connected to the first floor by a stairwell. The second story walls were wood framed with 51~mm (2~in) by 102~mm (4~in) studs set to 0.40~m (16~in) centers. The interior walls were protected by 16~mm (0.625~in) fire rated gypsum board, 16~mm (0.625~in) durarock board, and a second layer of 16~mm (0.625~in) fire rated gypsum board. The exterior walls were protected with 11~mm (0.44~in) oriented strand board and 8~mm (0.31~in) fiber cement lap siding.

\textit{Should the english units be in fractions instead of decimals? For example, 0.625 --> 5/8. And if decimals, should they all be rounded to same place?}

\subsubsection{Layout}
\label{sec:layout}
Dimensioned floor plans of the East and West Structures are presented in Figures~\ref{fig:east_dimensioned_plan}~and~\ref{fig:west_dimensioned_plan}, respectively.

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/East_Test_Structure_Dimensioned_Full}
	\caption[Dimensioned floor plan of the East Structure.]{Dimensioned floor plan of the East Structure. Interior dimensions are symmetric across horizontal and vertical centerlines.}
	\label{fig:east_dimensioned_plan}
\end{figure}

@@

\subsection{Instrumentation}
\label{sec:instrumentation}
A discussion of uncertainties for each type of measurement in this report can be found in Section~\ref{sec:uncertainty}. For every experiment, gas velocity at an interior doorway was measured using differential pressure transducers connected to bi-directional velocity probes~\cite{McCaffrey:Combustion_and_Flame}. An array of six bi-directional probes located at the northeast interior doorway of the East Structure is pictured below in Fig.~\ref{fig:BDPs}. A single thermocouple was attached to each bi-directional probes to measure the temperature to be used in conjunction with the pressure measurement to derive velocity. Every thermocouple attached to the bi-directional probes was exposed-bead, Chromel-Alumel (type K) with a 1.0~mm (0.04 in) diameter. Starting with the exposed bead, the thermocouple wire was sheathed in a 3.2~mm (0.13 in) diameter Inconel shield that was 0.76~m (2.5~ft) in length.

The approximate locations of the bi-directional probes used to measure air velocity are annotated in the plan views of the East Structure and West Structure experimental setups (Fig.~\ref{fig:east_setup}--\ref{fig:flow_path_2}) found in Section~\ref{sec:exp_procedure}. To minimize potential measurement errors due to external environmental conditions (i.e., wind), only velocity data measured by bi-directional probes located at interior doorways are considered in this report. The set of bi-directional probes at the northeast interior doorway of the East Structure, denoted as ``A6", contained six probes located at distances of 0.30~m (1.0~ft), 0.60~m (2.0~ft), 0.90~m (3.0~ft), 1.20~m (3.9~ft), 1.50~m (4.9~ft), and 1.80~m (5.9~ft) below the soffit of the doorway. The set of bi-directional probes at the interior doorway in the West Structure, denoted as ``A10", contained eight probes located at distances of 0.08~m (0.25~ft), 0.34~m (1.1~ft), 0.61~m (2.0~ft), 0.88~m (2.9~ft), 1.15~m (3.7~ft), 1.42~m (4.7~ft), 1.68~m (5.5~ft), and 1.95~m (6.4~ft) below the soffit of the corresponding doorway.

\begin{figure}[!ht]
	\includegraphics[width=3.5in]{../Figures/Pictures/BDPs_east}
	\caption[Array of bi-directional probes at interior doorway in East Structure.]{Array of bi-directional probes located at the northeast interior doorway in the East Structure. View is looking at the north side of the doorway.}
	\label{fig:BDPs}
\end{figure}
\FloatBarrier

@@

Water flow rate was measured with a pressure and flow meter combination shown in Fig.~\ref{fig:flow_meter}. The meter consists of a section of 6.35~cm (2.5~in) cast aluminum pipe with a 0--4.1 MPa (0--600 psi) pressure transducer and a paddlewheel type flow sensor with a range of 0 to 4800~lpm (1250~gpm). The pressure transducer and paddlewheel both connect to the battery operated control box where the pressure transducer voltage is converted to a pressure and the paddlewheel pulse count is converted to a volumetric flow rate. The manufacturer reports a $\pm5$\% calibration expanded uncertainty for the flow sensor and $\pm3$\%  for the pressure sensor~\cite{Akron}. The pressure transducer was calibrated with a known analog pressure gauge. The flow meter was calibrated by capturing water over time and measuring that mass of water to determine the flow rate. The total expanded uncertainty was estimated at $\pm10$\%. 
% FIND FLOW METER REFERENCE
\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Pictures/flow_meter}
	\caption[Flow meter used to measure flow rate during experiments.]{Pressure and flow meter combination used to measure water flow rate during the experiments.}
	\label{fig:flow_meter}
\end{figure}
\FloatBarrier

\section{Experimental Procedure}
\label{sec:exp_procedure}
Every experiment discussed in this report involved discharging water at a set flow rate from a combination nozzle in a straight, narrow fog, or wide fog stream pattern or from a smooth bore nozzle in a solid stream pattern. Specific doors were opened or closed before and/or during the experiments to change the ventilation in the structure.

A total of seven tests were conducted to study the impact of different hose stream and application patterns on gas movement in a structure; three used a combination nozzle attached to a monitor to flow water at approximately 120~GPM, three used a combination nozzle attached directly to a hoseline to flow water at approximately 120~GPM (Fig.~\ref{fig:monitor+handline}), and one used combination and smooth bore nozzles attached to a monitor to flow water at approximately 180~GPM. Experiments that used a monitor to flow water always applied the stream in a fixed position and thus were focused on changing the location of water application in addition to the stream pattern, while the experiments that used a combination nozzle attached directly to a hoseline to flow water focused on varying the pattern in which water was applied in addition to the stream pattern.

\textit{Should I present flow rates in GPM only, or LPM (GPM)? What about the tables? As of now, GPM is used, but that goes against the rule of keeping the type of units consistent in a table.}
\begin{figure}[!ht]
	\minipage{3in}
	\begin{center}
		\includegraphics[width=2.9in]{../Figures/Pictures/monitor}
	\end{center} 
	\endminipage \hfill
	\minipage{3in}
	\begin{center}
		\includegraphics[width=2.9in]{../Figures/Pictures/handline}
	\end{center}
	\endminipage \hfill
	\caption[Monitor and handline equipped with a combination nozzle.]{Monitor (left) and 1.75~in handline (right) equipped with a combination nozzle and used to flow water during experiments.}
	\label{fig:monitor+handline}
\end{figure}
\FloatBarrier

Three different structure layouts were used for the seven tests. The tests conducted in the East Structure, Tests 33 and 34, were performed using an identical structure layout, presented in Fig.~\ref{fig:east_setup}. Test 33 used a monitor and combination nozzle to flow water, and Test 34 used a combination nozzle attached directly to a handline to flow water.

The main difference between the two layouts used for the West Structure experiments is the position of the south side door on the first level. Tests 16 and 18 used the configuration with the door opened (Fig.~\ref{fig:flow_path_1}), and Tests 17, 19, and 70 used the configuration with the door closed (Fig.~\ref{fig:flow_path_2}). Tests 16 and 17 used a combination nozzle attached to a monitor to flow water; Tests 18 and 19 used a combination nozzle attached directly to a handline to flow water; and Test 70 used a combination nozzle and a smooth bore nozzle attached to a monitor to flow water in a straight and solid stream, respectively.

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/Specific_Tests/East_Hose_Test_Annotated}
	\caption[Plan view of the East Structure setup for Tests 33 and 34.]{Plan view of the East Structure setup for Tests 33 and 34. The view is annotated with the approximate location of water flow (nozzle graphic), the direction of the established flow path (red line), and the approximate location of bi-directional probe array A6 (blue square).}
	\label{fig:east_setup}
\end{figure}
\FloatBarrier

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/Specific_Tests/West_Hose_Test_2nd_Floor_Annotated}
	\\~\\
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/Specific_Tests/West_Hose_Test_18_1st_Floor_Annotated}
	\caption[Plan view of the West Structure setup for Tests 16 and 18.]{Plan view of the second floor (top) and first floor (bottom) setup in the West Structure for Tests 16 and 18. The approximate location of water flow (nozzle graphic), the direction of the established flow path (red line), and the approximate location of bi-directional probe array A10 (blue square). The stairwell door and north side, west double door on the second floor were opened and closed at certain instances during the tests; the flow path was considered fully established when both doors were in the opened position.}
	\label{fig:flow_path_1}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/Specific_Tests/West_Hose_Test_2nd_Floor_Annotated}
	\\~\\
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/Specific_Tests/West_Hose_Test_19_1st_Floor_Annotated}
	\caption[Plan view of the West Structure setup for Tests 17 and 19.]{Plan view of the second floor (top) and first floor (bottom) setup in the West Structure for Tests 17 and 19. The approximate location of water flow (nozzle graphic), the direction of the established flow path (red line), and the approximate location of bi-directional probe array A10 (blue square). The stairwell door and north side, west double door on the second floor were opened and closed at certain instances during the tests; the flow path was considered fully established when both doors were in the opened position.}
	\label{fig:flow_path_2}
\end{figure}
\FloatBarrier

\subsection{Monitor Experiments}
\label{sec:monitor_procedure}
Three different tests were conducted using a monitor equipped with a combination nozzle to flow water at 120~GPM in a straight stream, narrow fog stream, and wide fog stream pattern. One test, Test 33, occurred in the East Structure, and the two other tests, Tests 16 and 17, occurred in the West Structure. A fourth test, Test 70, used a combination nozzle and a smooth bore nozzle attached to a monitor to flow water at 180~GPM in a straight stream and solid stream, respectively. Because monitor nozzles are designed to flow water from a set position, water was always applied in a fixed position during the monitor experiments. Thus, the monitor experiments primarily focused on the impact of different hose stream patterns on air movement in a structure.

Test 33 utilized a monitor equipped with a combination nozzle to flow water from Room C of the East Structure to the area above the doorway on the south wall of Room B (Fig.~\ref{fig:test_33_pic}). A plan view of the structure's layout annotated with the approximate location of the bi-directional probes array used to measure gas velocity through an interior doorway (A6) and the general direction of the fully established flow path for Test 33 can be found in Fig.~\ref{fig:east_setup}. The procedure for Test 33 was the simplest of all the test procedures. First, water flowed in a straight stream pattern for 30 seconds, then the stream was changed to a narrow fog for 30 seconds, and finally to a wide fog stream for 30 seconds. The procedure was repeated an additional time.
